---
title: "Infill-Analysis"
date: "2023-08-15"
Author: Thomas Cantu
output: word_document
---
```{r echo=FALSE, include=FALSE  }
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(dplyr)
library(tidyverse)
library(forecast)
library(zoo) 
library(readxl)
```


The data analysis presented here was gathered through a collaboration with the El Paso Economic Development and International Business Department. The dataset comprises historical infill agreement projects aligned with the Infill Development Incentive Program, which received approval from the city council in 2010. The information is derived from the Master Agreements document, which is diligently maintained by the Compliance Sector within the Economic Development and International Business Department.

# The Data 
In this study, we have harnessed the valuable data made available through the cooperation of the El Paso Economic Development and International Business Department. The dataset encompasses historical records of infill agreement projects closely tied to the Infill Development Incentive Program, an initiative that received endorsement from the city council back in 2010. To ensure accuracy and consistency, the data is sourced from the meticulously maintained Master Agreements document, overseen by the diligent team at the Compliance Sector within the Economic Development and International Business Department.
```{r , echo=FALSE}

df <- read.csv("C:/Users/CantuCT/Documents/Infill Analysis/filtered_infill_data.csv")
df2 <- read.csv("C:/Users/CantuCT/Documents/Infill Analysis/2023 Complete projects Infill taxes.csv")
# C:\Users\CantuCT\Documents\Infill Analysis
# C:/Users/cantu/OneDrive - Bellevue University/Documents/City El Paso/filtered_infill_data.csv
```

```{r, echo=FALSE}
 #head(df[, 2:4], n=3)
```

# Total Incentives Paid To Date 

The figure presented represents the cumulative total of incentives disbursed to date for all the projects approved under the Infill Development Incentive Program by the City Council. This sum encompasses various forms of incentives, including but not limited to tax rebates, grants, and construction fee waivers, as stipulated in the infill policy for eligible applicants. It reflects the comprehensive support provided to qualified participants in their efforts to promote infill development within the city.

```{r, echo=FALSE}
#df$Total_Incentives_paid_to_date <- as.numeric
#print(df$Total_Incentives_paid_to_date)
total_sum <- sum(df$Total_Incentives_paid_to_date, na.rm = TRUE)
cat("Total Sum of Incentives Paid to Date:", '$',total_sum, "\n")
```
# Project Status 

The following analysis provides an overview of the status of infill projects from 2012 to 2023. These projects are classified into four categories:

**Active:** These are projects that are currently ongoing and fall within the guidelines of the current infill program, having received approval from the City Council. These projects are actively progressing towards their completion and continue to benefit from the incentives provided.

**Terminated:** This category includes projects that were terminated by the Economic and International Business Department for various reasons. The termination could be due to non-compliance, failure to meet program requirements, or other factors leading to the discontinuation of the project.

**Closed:** Projects falling into this category are those that forfeited their incentives, which could include tax rebates, grants, or other incentives offered under the infill program. The reasons for forfeiture could range from non-compliance with program terms to voluntary withdrawal from the incentives.

**Withdrawn:** Projects in this classification are those where the recipients of infill incentives expressed their desire to be removed from the program. This decision could arise from changes in project priorities or other strategic considerations.

By examining the data on projects within these categories, we gain valuable insights into the effectiveness and outcomes of the infill program over the specified period, helping us understand the dynamics of infill development in the city.


```{r, echo=FALSE}
df$Subtype_Status_of_the_Project <- trimws(df$Subtype_Status_of_the_Project)
unique_status_subtypes <- table(df$Subtype_Status_of_the_Project)

cat("\nStatus Subtypes:\n")
for (subtype in names(unique_status_subtypes)) {
  count <- unique_status_subtypes[subtype]
  formatted_output <- sprintf("%-40s %d", subtype, count)
  cat(formatted_output, "\n")
}

```


```{r, echo=FALSE, dpi = 300}
value_counts <- table(df$Subtype_Status_of_the_Project)

# Create a bar plot
status_plot <- ggplot(data = data.frame(Category = names(value_counts), Count = as.numeric(value_counts)),
            aes(x = reorder(Category, -Count), y = Count, fill = Category)) +
  geom_bar(stat = "identity", width = 0.8) +
  geom_text(aes(label = Count), vjust = -0.3, color = "black") +  # Add count labels
  labs(x = "Category", y = "Count", title = "Status of the Projects from 2012 to 2023") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0, max(value_counts) + 2))  # Adjust y-axis limit


# show the plot 
print(status_plot)
```

### Percentages 


```{r, echo=FALSE, dpi = 300}

# Calculate percentages
percentages <- prop.table(table(df$Subtype_Status_of_the_Project)) * 100

# Round percentages to one decimal place
rounded_percentages <- round(percentages, 1)

# Create a data frame with rounded percentages
percentage_df <- data.frame(Category = names(rounded_percentages),
                            Percentage = as.numeric(rounded_percentages))

# Create a horizontal stacked bar chart with percentages
status_percent <- ggplot(percentage_df, aes(x = Category, y = Percentage, fill = Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(Percentage, "%")), 
            position = position_stack(vjust = 0.5), 
            size = 3) +
  coord_flip() +
  labs(title = "Percentages of Project Categories", x = NULL, y = "Percentage") +
  theme_minimal() +
  theme(legend.position = "none")

# Show the plot
print(status_percent)

```
```{r, echo=FALSE}
total_count_projects <- sum(value_counts)
non_active_projects <- total_count_projects - 13  # 13 projects are active
sucess_rate <- (26/total_count_projects) * 100  # 18 projects have been completed, 13 are active # update -5 no incentives were paid out =26

cat('\nTotal Projects:', total_count_projects, '\n')
cat('Total Non-Active Projects:', non_active_projects, '\n')
cat('Success Rate of the Infill Program:', sucess_rate, '%\n')
```
## Status Insights

### The primary factors contributing to the closure, termination, or withdrawal of projects can be categorized as follows:

**Projects that never started construction:** These are projects that were conceptualized but never initiated the construction phase. The reasons for this could include insufficient funding, changes in market conditions, or regulatory hurdles that prevented the project from moving forward.

**Projects that did not finish construction of the property:** These are projects that started construction but were not completed for various reasons. Possible factors could include budgetary constraints, unexpected challenges during construction, or legal disputes that halted progress.

**Projects that forfeited tax rebates:** Some projects may have initially qualified for tax rebates or incentives but later had to forfeit them due to non-compliance with certain regulations or failure to meet specified project milestones.

### Hypotheses related to these contributors are as follows:

1. Projects obtaining infill agreements may not be fully planned and developed at the time of obtaining the agreement. 

2. The infill incentive, intended to encourage urban development, might not align well with projects in their early stages of development. As a result, the project owners might face challenges in utilizing the incentive effectively.

3. The infill process itself may prove to be overly complex for project owners to navigate, leading to delays, confusion, and potential complications during the project's execution.

4. In some cases, the incentives provided through the infill policy might not be sufficient or attractive enough to adequately benefit project owners, discouraging them from taking advantage of the program.

Understanding and addressing these factors can help improve project success rates and ensure that the benefits of infill policies and incentives are maximized for both developers and the community.

# Projects Over Time 

```{r, echo=FALSE, dpi = 300}
# Get the counts of projects by fiscal year
project_counts_by_year <- table(df$Fiscal_Year)

# Convert the table to a data frame
project_counts_df <- as.data.frame(project_counts_by_year)

# Plot the count of projects by fiscal year using ggplot2
ggplot(data = project_counts_df, aes(x = Var1, y = Freq)) +
  geom_bar(stat = "identity", fill = "darkblue") +
  geom_text(aes(label = Freq), vjust = -0.5, color = "black", size = 3) +  # Add count labels
  labs(x = "Year", y = "Count", title = "Number of Projects by Year") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_cartesian(ylim = c(0, max(project_counts_df$Freq) + 5)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(axis.title = element_text(size = 12, face = "bold"))
```


# Potential Projects To 2025

Presented below is a projected forecast of potential new projects until the year 2027, based on an analysis of historical project patterns over time. The forecast indicates potential outcomes for projects if the Infill Development Incentive Program were to continue at its current pace, maintaining its existing structure.

By extrapolating data from past project trends, we observe a potential trend of continued growth in the number of projects under the Infill Development Incentive Program. The program's current incentives and framework are projected to play a role in encouraging infill development and potentially attracting more participants.

The forecast highlights the possible impact and effectiveness of the Infill Development Incentive Program in contributing to economic growth and urban revitalization within the city. Should the program maintain its current trajectory, these projections suggest that the number of infill projects may increase over the specified period, contributing to the ongoing development of the community.

```{r, echo=FALSE, dpi = 300}

# Get the counts of projects in each fiscal year
project_counts_by_year <- table(df$Fiscal_Year)

# Convert the table to a data frame
project_counts_df <- as.data.frame(project_counts_by_year)

# Calculate the rolling mean (simple moving average) with a window of 3
rolling_mean_window <- 3
rolling_mean <- rollmean(project_counts_df$Freq, rolling_mean_window, fill = NA)

# Create a time series object
project_counts_ts <- ts(project_counts_df$Freq)

# Forecast the next 5 years based on the rolling mean using the forecast package
forecast_values <- forecast::forecast(auto.arima(project_counts_ts), h = 5)

# Create a data frame with the forecasted values
forecast_df <- data.frame(Year = seq(max(as.numeric(names(project_counts_by_year))) + 1, length.out = 5), Forecasted_Projects = forecast_values$mean)

# Plot the original data and the forecasted values using ggplot2
ggplot() +
  geom_line(data = project_counts_df, aes(x = as.numeric(as.character(Var1)), y = Freq), color = 'darkblue') +
  geom_line(data = forecast_df, aes(x = Year, y = Forecasted_Projects), color = 'red') +
  labs(x = "Year", y = "Number of Projects", title = "Number of Projects by Year (Including Forecast to 2027)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(axis.title = element_text(size = 12, face = "bold")) +
  scale_color_manual(values = c('darkblue', 'red'), labels = c('Original Data', 'Forecasted Data')) +
  guides(color = guide_legend(title = "Data"))
```
```{r, include=FALSE }
# Get the counts of projects in each fiscal year
project_counts_by_year <- table(df$Fiscal_Year)

# Convert the table to a time series object
project_counts_ts <- ts(project_counts_by_year, start = min(df$Fiscal_Year), frequency = 1)

# Fit an exponential smoothing state space model (ETS) to the data
ets_model <- ets(project_counts_ts)

# Forecast the next 5 years using the ETS model
forecast_values <- forecast(ets_model, h = 5)

# Create a data frame with the forecasted values
forecast_df <- data.frame(Year = seq(max(df$Fiscal_Year) + 1, length.out = 5), Forecasted_Projects = forecast_values$mean)

# Plot the original data and the forecasted values using ggplot2
ggplot() +
  geom_line(data = data.frame(Year = as.numeric(names(project_counts_by_year)), Original_Data = as.numeric(project_counts_by_year)), aes(x = Year, y = Original_Data), color = 'darkblue') +
  geom_line(data = forecast_df, aes(x = Year, y = Forecasted_Projects), color = 'red') +
  labs(x = "Year", y = "Number of Projects", title = "Number of Projects by Year (Including Forecast)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(axis.title = element_text(size = 12, face = "bold")) +
  scale_color_manual(values = c('darkblue', 'red'), labels = c('Original Data', 'Forecasted Data')) +
  guides(color = guide_legend(title = "Data"))
```

# Projects by Industry 
```{r, echo=FALSE}
df$Industry_Sector <- trimws(df$Industry_Sector)
unique_industry_subtypes <- table(df$Industry_Sector)
cat("\nIndustry Sector:\n")
print(unique_industry_subtypes)

```
```{r, echo=FALSE, dpi = 300}


# Create a bar plot for industries 
status_plot_industries <- ggplot(data = data.frame(Category = names(unique_industry_subtypes), Count = as.numeric(unique_industry_subtypes)),
                                  aes(x = reorder(Category, -Count), y = Count, fill = Category)) +
  geom_bar(stat = "identity", width = 0.8) +
  geom_text(aes(label = Count), vjust = -0.5, color = "black", size = 3) +
  labs(x = "Category", y = "Count", title = "Industries Sectors of Infill Projects") +
  scale_fill_brewer(palette = "Set1") +  # Set the color palette
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0, max(as.numeric(unique_industry_subtypes)) + 2))

# show the plot 
print(status_plot_industries)
```


# Notice 

The following information presented has been obtained from the El Paso Central Appraisal District (EPCAD). EPCAD has provided a disclosure regarding the accuracy of the data.

**El Paso Central Appraisal District Disclosure** 

**All information contained herein, is considered in the public domain and is distributed without warranty of any kind, implied, expressed or statutory. The El Paso Central Appraisal District makes no claims, promises or guarantees about the accuracy, completeness, or adequacy of this information and expressly disclaims liability for any errors and omissions. Information relating to 2024 should be considered a ''work in progress''. Prior year data is informational only and does not necessarily reflect the values certified to the tax office. Building and land detail information may not be available prior to 2013.**




# Property Values for Completed Projects  

## Property Value Increase

The data provided illustrates the difference between the initial property value at the start of the infill agreement and its value when the agreement concluded. It's worth emphasizing that not all property valuations were attainable through EPCAD. Consequently, properties with unavailable data were omitted from the mean calculation. Additionally, a distribution of property value increases by percentages has been included. These values were retained, guided by the assumption that the absent data points might correspond to outliers in the present dataset.

```{r, include=FALSE}
df$Property_Value_at_Exe.<- as.numeric(df$Property_Value_at_Exe.) # turn variables to numeric 
df$Property_Value_at_Exp<- as.numeric(df$Property_Value_at_Exp)

```

```{r, echo=FALSE}
#Property_value_change <- (df$Property_Value_at_Exp - df$Property_Value_at_Exe.)
#print(Property_value_change)
```
```{r, echo=FALSE}
completed_projects <- df[df$Subtype_Status_of_the_Project == 'Completed', ] # new dataframe of only completed projects 
#print(completed_projects)
#completed_Df <- completed_projects

```

```{r, echo=FALSE}

Property_value_change <- ifelse(completed_projects$Property_Value_at_Exp == 0 | completed_projects$Property_Value_at_Exe == 0, 
                                 NA, 
                                 completed_projects$Property_Value_at_Exp - completed_projects$Property_Value_at_Exe)

completed_projects$Property_value_change <- ifelse(completed_projects$Property_Value_at_Exp == 0 | completed_projects$Property_Value_at_Exe == 0, 
                                                    NA, 
                                                    completed_projects$Property_Value_at_Exp - completed_projects$Property_Value_at_Exe)

completed_projects$Property_value_change <- as.numeric(completed_projects$Property_value_change)

formatted_property_value_change <- format(completed_projects$Property_value_change, big.mark = ",", prefix = "$") # format to look pretty

#cat(formatted_property_value_change, "\n") # Print the formatted property value changes



```

```{r, echo=FALSE}
sum_property_change <- sum(Property_value_change, na.rm = TRUE) # skip NAs sum of all the property values that have increased 
non_NA <- sum(!is.na(Property_value_change)) # number of property skip NAs
#print(non_NA)
average_property_change <- sum_property_change/non_NA # take the average 
formatted_average <- format(average_property_change, big.mark = ",") # add some comas to be pretty 
cat("The average increase in property value from execution of agreements to completion is", "$", formatted_average, "\n")
```

```{r, echo=FALSE, dpi = 300}

# Calculate property value change as percentage change
Property_value_change_percentage <- ifelse(completed_projects$Property_Value_at_Exp == 0 | completed_projects$Property_Value_at_Exe == 0, 
                                                    NA, 
                                                    ((completed_projects$Property_Value_at_Exp - completed_projects$Property_Value_at_Exe) / completed_projects$Property_Value_at_Exe) * 100)

completed_projects$Property_value_change_percentage <- ifelse(completed_projects$Property_Value_at_Exp == 0 | completed_projects$Property_Value_at_Exe == 0, 
                                                    NA, 
                                                    ((completed_projects$Property_Value_at_Exp - completed_projects$Property_Value_at_Exe) / completed_projects$Property_Value_at_Exe) * 100)

#print(Property_value_change_percentage)

# Create a histogram plot of Property_value_change_percentage, skipping NAs
ggplot(data = completed_projects[!is.na(completed_projects$Property_value_change_percentage), ], 
       aes(x = Property_value_change_percentage)) +
  geom_histogram(binwidth = 15, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Property Value Change Percentage",
       x = "Percentage Change",
       y = "Frequency")

```







```{r, echo=FALSE}
## find years of agreement 
# Convert the "Exp._Date" column to a Date format
df2$Exp._Date <- as.Date(df2$Exp._Date, format = "%m/%d/%Y %H:%M")
df2$Exe._Date <- as.Date(df2$Exe._Date, format = "%m/%d/%Y %H:%M")
# Calculate the number of years
years_diff <- as.numeric(difftime(df2$Exp._Date, df2$Exe._Date, units = "days") / 365.25)

# Round the years_diff to the nearest whole number
years_diff <- round(years_diff)

df2 <- cbind(df2[, 1:7], years_diff, df2[, 8:length(df2)])

# Print the updated dataframe
#print(df2)



```


```{r,echo=FALSE}
# new columns for ROI 
#Create a sequence of years from 2025 to 2048 
years <- seq(2025,2048, by = 1)
# new columns loop
for (year in years){
  df2[paste("Pvalue_", year, sep = "")] <- NA
  df2[paste("Tvalue_", year, sep = "")] <- NA
  df2[paste("County_", year, sep = "")] <- NA
}

# print(df2) 
```


```{r, echo=FALSE}
# setting up 2024 to run an easy loop for all future  years 
# Calculate Pvalue_2024 as Pvalue_2023 * 1.03, but only if Pvalue_2024 is NA

df2$Pvalue_2024 <- ifelse(is.na(df2$Pvalue_2024), round(df2$Pvalue_2023 * 1.03), df2$Pvalue_2024)

# Check if Tvalue_2024 and County_2024 are NA
missing_tvalue <- is.na(df2$Tvalue_2024)
missing_county <- is.na(df2$County_2024)

# Calculate Tvalue_2024 and County_2024 only for rows where they are NA
df2$Tvalue_2024[missing_tvalue] <- round(df2$Pvalue_2024[missing_tvalue] * 0.00818875, 2)
df2$County_2024[missing_county] <- round(df2$Pvalue_2024[missing_county] * 0.00426289, 2)

# Print the updated dataframe
#print(df2)



```

```{r,echo=FALSE}
# test 1 to try to so all years 

#years_already_have  <- 2024 - as.integer(format(df2$Exp._Date, "%Y"))
#print(years_already_have)

#years_remaining <- 25 - years_already_have
#print(years_remaining)
```

```{r, echo=FALSE}
# Calculate years_already_have for each row
years_already_have <- 2024 - as.integer(format(df2$Exp._Date, "%Y"))

# Calculate years_remaining for each row
years_remaining <- 25 - years_already_have

# Loop through rows
for (row in 1:nrow(df2)) {
  # Loop to update columns from 2025 to 2048
  for (i in 2025:2048) {
    # Create column names
    pvalue_col <- paste0("Pvalue_", i)
    tvalue_col <- paste0("Tvalue_", i)
    county_col <- paste0("County_", i)

    # Get the years_remaining for the current row
    current_years_remaining <- years_remaining[row]

    # Check if current_years_remaining is zero, and if so, fill columns with zeros and break
    if (current_years_remaining <= 0) {
      df2[row, (pvalue_col)] <- 0
      df2[row, (tvalue_col)] <- 0
      df2[row, (county_col)] <- 0
      break
    }

    # Calculate Pvalue for the current year
    df2[row, (pvalue_col)] <- ifelse(
      is.na(df2[row, (pvalue_col)]),
      df2[row, paste0("Pvalue_", i - 1)] * 1.03,
      df2[row, (pvalue_col)]
    )

    # Check if Tvalue and County are NA
    missing_tvalue <- is.na(df2[row, (tvalue_col)])
    missing_county <- is.na(df2[row, (county_col)])

    # Calculate Tvalue and County only for rows where they are NA
    df2[row, (tvalue_col)][missing_tvalue] <- df2[row, (pvalue_col)][missing_tvalue] * 0.00818875
    df2[row, (county_col)][missing_county] <- df2[row, (pvalue_col)][missing_county] * 0.00426289

    # Decrease current_years_remaining
    current_years_remaining <- current_years_remaining - 1

    # Ensure current_years_remaining is not negative
    current_years_remaining <- max(0, current_years_remaining)

    # Update years_remaining for the current row
    years_remaining[row] <- current_years_remaining
  }
}

# Print the updated dataframe
#print(df2)



```

```{r, echo=FALSE}
# vectors to search key words
tvalue_columns <- grep("Tvalue_", names(df2), value = TRUE)
county_columns <- grep("County_", names(df2), value = TRUE)

#calculate the sums for each row

df2$Total_Tvalue <- rowSums(df2[, tvalue_columns], na.rm = TRUE)
df2$Total_County <- rowSums(df2[, county_columns], na.rm = TRUE)

#print(df2)

```



## Property Value Average for Completed Projects  

```{r, echo=FALSE}
# calculate avg property value in 2024
avg_Pvalue_2024 <- mean(df2$Pvalue_2024, na.rm = TRUE)
avg_Pvalue_2024<- format(avg_Pvalue_2024, big.mark = ",") # add some comas to be pretty 
cat("The average property value in 2024 is", "$", avg_Pvalue_2024, "\n")

# calculate avg property value in 2045
avg_Pvalue_2045 <- mean(df2$Pvalue_2045, na.rm = TRUE)
avg_Pvalue_2045 <- format(avg_Pvalue_2045, big.mark = ",") # add some comas to be pretty 
cat("\nThe average property value in 2045 is", "$", avg_Pvalue_2045, "\n")


```

# City and County Benifits

## Singular Benifits

Below, you will find the average tax benefits for both the city and county. These tax benefits have been obtained directly from EPCAD (El Paso Central Appraisal District). The process of gathering this property tax information involved collecting data for the year 2024 and 2045, followed by a meticulous calculation of the mean. It's noteworthy that missing data points were excluded from the calculation; however, we made a deliberate choice to include potential outliers, considering that missing data might correlate with outlier values. This comprehensive approach ensures a robust representation of the average tax benefits for the city and county.


### City
```{r, echo=FALSE}
# calculate avg property tax for city in 2024
avg_Tvalue_2024 <- mean(df2$Tvalue_2024, na.rm = TRUE)
avg_Tvalue_2024<- format(avg_Tvalue_2024, big.mark = ",") # add some comas to be pretty 
cat("The average city property tax benefit in 2024 is", "$", avg_Tvalue_2024, "\n")


# calculate avg property tax for city  in 2045
avg_Tvalue_2045 <- mean(df2$Tvalue_2045, na.rm = TRUE)
avg_Tvalue_2045 <- format(avg_Tvalue_2045, big.mark = ",") # add some comas to be pretty 
cat("\nThe average city property tax benefit in 2045 is", "$", avg_Tvalue_2045, "\n")


```
### County 
```{r, echo=FALSE}
# calculate avg property tax for county  in 2024
avg_County_2024 <- mean(df2$County_2024, na.rm = TRUE)
avg_County_2024<- format(avg_County_2024, big.mark = ",") # add some comas to be pretty 
cat("\nThe average county tax benefit in 2024 is", "$", avg_County_2024, "\n")

# calculate avg property tax for county in 2045
avg_County_2045 <- mean(df2$County_2045, na.rm = TRUE)
avg_County_2045 <- format(avg_County_2045, big.mark = ",") # add some comas to be pretty 
cat("\nThe average county tax benefit in 2045 is", "$", avg_County_2045, "\n")


```



## Total Net Benefits

The figures presented represent the comprehensive net benefits accrued from all qualifying infill projects upon the culmination of a 25-year analysis encompassing both the city and county of El Paso.

These net benefits have been calculated by considering a consistent annual property value appreciation rate of 3% throughout the entire analysis period. Furthermore, they account for the respective tax rates applicable to both the city and county, which stand at 0.818875 and 0.426289, respectively.


```{r, echo=FALSE}
# For city tax benefit in 2024 # no need to include have it just in case 

# Specify the column names for Tvalue_2019 to Tvalue_2024
tvalue_cols <- c("Tvalue_2019", "Tvalue_2020", "Tvalue_2021", "Tvalue_2022", "Tvalue_2023", "Tvalue_2024")

# Calculate the sums of the specified columns row-wise, skipping NA values
sums <- rowSums(df2[, tvalue_cols], na.rm = TRUE)

# Calculate the total sum of the sums
total_sum_tvalue <- sum(sums, na.rm = TRUE)
total_sum_tvalue <- format(total_sum_tvalue, big.mark = ",")

# Calculate the average of the sums
average_sum_tvalue <- mean(sums, na.rm = TRUE)
average_sum_tvalue <- format(average_sum_tvalue, big.mark = ",")

# Print the total sum and average of Tvalue columns from 2019 to 2024 for all rows
#cat("Total city net benefit for 2024: $", total_sum_tvalue, "\n") # no need to include
#cat("\nAverage city net benefit per project for 2024: $", average_sum_tvalue, "\n") # no need to include 

```



```{r, echo=FALSE}
# For county tax benefit in 2024 # no need to include have it just in case 

# Specify the column names for County_2019 to County_2024
county_cols <- c("County_2019", "County_2020", "County_2021", "County_2022", "County_2023", "County_2024")

# Calculate the sums of the specified columns row-wise, skipping NA values
sums_county <- rowSums(df2[, county_cols], na.rm = TRUE)

# Calculate the total sum of the sums
total_sum_county <- sum(sums_county, na.rm = TRUE)
total_sum_county <- format(total_sum_county, big.mark = ",")

# Calculate the average of the sums
average_sum_county <- mean(sums_county, na.rm = TRUE)
average_sum_county <- format(average_sum_county, big.mark = ",")

# Print the total sum and average of County columns from 2019 to 2024 for all rows
#cat("Total county net benefit for 2024: $", total_sum_county, "\n")
#cat("\nAverage county net benefit per project for 2024: $", average_sum_county, "\n")

```

### City
```{r, echo=FALSE}
# total city benefit  investment at 30 year investment period 
# Calculate the sum of Total_Tvalue, skipping NA values
total_sum_total_tvalue <- sum(df2$Total_Tvalue, na.rm = TRUE)
total_sum_total_tvalue1 <- format(total_sum_total_tvalue, big.mark = ",") # add some commas to be pretty

# Calculate the average of Total_Tvalue, skipping NA values
average_total_tvalue <- mean(df2$Total_Tvalue, na.rm = TRUE)
average_total_tvalue <- format(average_total_tvalue, big.mark = ",") # add some commas to be pretty

# Print the total sum and average of Total_Tvalue
cat("The total city benefits for all Infill completed projects at 25 years: $", total_sum_total_tvalue1, "\n")
cat("\nAverage city benefit for a completed project at 25 years: $", average_total_tvalue, "\n")


```
### County
```{r, echo=FALSE}
# Total county benefit of infill completed projects over a 30-year investment period
# Calculate the sum of Total_County, skipping NA values
total_sum_total_county <- sum(df2$Total_County, na.rm = TRUE)
total_sum_total_county <- format(total_sum_total_county, big.mark = ",") # add some commas to be pretty

# Calculate the average of Total_County, skipping NA values
average_total_county <- mean(df2$Total_County, na.rm = TRUE)
average_total_county <- format(average_total_county, big.mark = ",") # add some commas to be pretty

# Print the total sum and average of Total_County
cat("The total county benefits for all infill completed projects at 25 years: $", total_sum_total_county, "\n")
cat("\nAverage county benefit for a  completed project at 25 years: $", average_total_county, "\n")

```
## ROI

The figure presented illustrates the Return on Investment (ROI) percentages for the Economic Developments Infill Program. This ROI metric has been calculated over a substantial 25-year analysis period, commencing from the culmination of qualifying projects within the program.

To compute this ROI, a consistent annual property value increase of 3% has been assumed over the entire analysis duration. Additionally, it takes into account the prevailing tax rate for the city of El Paso, which stands at 0.818875.

This data provides valuable insights into the long-term financial benefits and viability of the Economic Developments Infill Program, offering a glimpse into the potential returns that investors, developers, and stakeholders can anticipate over a quarter-century timeframe. It also underscores the economic growth and stability that this program can bring to the city of El Paso, making it an important consideration for those involved in urban development and investment planning.

```{r, echo=FALSE}
# ROI


# Calculate ROI and store it in a new column
df2$ROI <- ((df2$Total_Tvalue - df2$Total_Incentives_paid_to_date) / df2$Total_Incentives_paid_to_date) * 100

# Print ROI values

#print(df2$ROI)

# sum total incentives paid out 
total_sum_Incentives_paid_to_date <- sum(df2$Total_Incentives_paid_to_date, na.rm = TRUE)
# Calculate the average return

average_return <- ((total_sum_total_tvalue - total_sum_Incentives_paid_to_date)/ total_sum_Incentives_paid_to_date) *100

# Print the average return
cat("The average rate of return on an Infill agreement is", average_return, "%." )
#print(average_return)

```
```{r, echo=FALSE}
#print(df2)
#file_name = "infill_ROI.csv"
#file_path = "C:/Users/CantuCT/Documents/Infill Analysis"
#write.csv(df2, file.path(file_path, file_name), row.names = FALSE)

```


